---
title: "R Workshop 3: Zelig"
author: "Dr. Ches Thurber"
date: "October 5, 2018"
output: 
  ioslides_presentation
---

## What's Zelig?

- A Woody Allen movie from 1983
- A package universe developed by Gary King at Harvard
- Another attempt to make R more like Stata
- Uniform implementation of regression models across model types
- Integrates many of King's statistical innovations (of which there are many)
- Implementation of "Clarify" i.e. post-estimation simulations
- can be fickle and quirky: *Zelig does NOT like missing data!*

##Election Data 
Let's load the data about presidential election outcomes and economic growth and check it out.

```{r Load Data, echo=T}
PresVote<-read.csv("PresVote.csv")
head(PresVote)
```

##Let's Make a Scatterplot
```{r Scatterplot, eval=T, echo=F}
plot(vote~rdi4, data=PresVote, pch=8, col="blue",
  xlab = "% Change in Income",
     ylab = "Incumbent Party's Vote %",
     main = "Economic Growth and Presidential Elections")
text(rdi4~vote, data=PresVote, labels=year, cex=.7, pos=1, offset=.5)
```

##Fitting a Linear Model in Base R
```{r Linear Model, echo=T}
Model1<-lm(vote~rdi4, data=PresVote)
summary(Model1)
```

##Put the Line on Our Plot
```{r abline Plot, eval=T, echo=F}
plot(vote~rdi4, data=PresVote, pch=8, col="blue",
  xlab = "% Change in Income",
     ylab = "Incumbent Party's Vote %",
     main = "Economic Growth and Presidential Elections")
text(vote~rdi4, data=PresVote, labels=year, cex=.7, pos=1, offset=.5)
abline(Model1, col="red")
```

##Running OLS in Zelig
```{r Zelig Model, echo=T, message=FALSE, warning=FALSE}
#install.packages("zeligverse")
library(zeligverse)
Z1<-zelig(vote~rdi4, model="ls", data=PresVote)
summary(Z1)
```

##Simulate Expected Values

```{r}
logrowth <- setx(Z1, rdi4 = 2)
higrowth <- setx(Z1, rdi4 = 4)
Sim1 <- sim(Z1, x = logrowth, x1 = higrowth)
summary(Sim1)
```

##Plot them
```{r}
par(mar = rep(2, 4))
plot(Sim1)
```

##Looking Under the Hood a Bit...
```{r}
SimOut<-zelig_qi_to_df(Sim1)
SlimOut<-qi_slimmer(SimOut)
```

## Simulate Continuously
```{r}
congrowth <- setx(Z1, rdi4 = seq(-2,4,by= .1))
Sim2 <- sim(Z1, x = congrowth)
plot(Sim2)
```

##Make that in GGPLOT
```{r message=FALSE, warning=FALSE, echo=F}
Sim2.Out<-qi_slimmer(zelig_qi_to_df(Sim2))
Sim2.Out$rdi4<-Sim2.Out$temp_fitted.u...
#install.packages("ggplot2")
library(ggplot2)
ggplot(Sim2.Out, aes(rdi4, qi_ci_median)) +
    geom_ribbon(aes(ymin = qi_ci_min, ymax = qi_ci_max), 
                alpha = 0.3) +
    geom_line() + 
    ylab('Expected Incumbent Vote Share') +
  ggtitle("Effect of Income Growth on Presidential Elections")+
    theme_bw()
```

